---
output:
  pdf_document: default
  html_document: default
---
PCA and Clustering on lettuce farm and retail prices

```{r message = FALSE, warning = FALSE}
# Loadpackages
library(fpp3)
library(tsibble)
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

```{r}
# read data with phoenix -> chicago reefer lane, trucking volume, and yuma weather
data_raw <- readr::read_csv(file = "data/data_phoenix_with_yuma_weather_and_volume_and_lags.csv")  %>%
  mutate(yw = yearweek(yw)) %>%
  select(-X1) %>%
  as_tsibble(key = c(Mode,ORegionDAT, DRegionDAT), index = yw) %>%
  relocate(yw, Mode, ORegionDAT, DRegionDAT, approx_cost, approx_vol, tmax, prcp)

data_raw
```

```{r, warning = FALSE, message = FALSE}
# Import lettuce data
lettuce <- readr::read_csv(file = 'data/lettuce_wholesale/all_lettuce.csv') %>%
  dplyr::select(-X1) %>%
  mutate(yw = yearweek(yw)) %>%
  relocate(commodity,yw)%>%
  as_tsibble(key = commodity,
             index = yw)

romaine <- lettuce %>% filter(commodity=="RO")
iceberg <- lettuce %>% filter(commodity=="IC")
green_leaf <- lettuce %>% filter(commodity=="GL")
red_leaf <- lettuce %>% filter(commodity=="RL")

print(head(romaine))
```

Visualize what the 4 lettuce price series look like.
```{r}
#visualize farm prices
autoplot(romaine, farm_price, color = "black") +
  autolayer(iceberg, farm_price, color="purple") +
  autolayer(green_leaf, farm_price, color="green") +
  autolayer(red_leaf, farm_price, color="red") +
  autolayer(data_raw %>% filter_index(~"2019 W26"), approx_cost, color="black", linetype="dotted") + 
  autolayer(data_raw %>% filter_index(~"2019 W26"), sanitized_cost, color ="black")

#visualize chciago retail prices
autoplot(romaine, chicago_retail, color = "black") +
  autolayer(iceberg, chicago_retail, color="purple") +
  autolayer(green_leaf, chicago_retail, color="green") +
  autolayer(red_leaf, chicago_retail, color="red")
```
Get data into right format for PCA and clustering
```{r}
romaine_clust <- as_tibble(romaine) %>%
  mutate(ro_farm_price = farm_price, ro_chicago_retail = chicago_retail) %>%
  select(yw, ro_farm_price, ro_chicago_retail) 

iceberg_clust <- as_tibble(iceberg) %>%
  mutate(ic_farm_price = farm_price, ic_chicago_retail = chicago_retail) %>%
  select(yw, ic_farm_price, ic_chicago_retail)

green_leaf_clust <- as_tibble(green_leaf) %>%
  mutate(gl_farm_price = farm_price, gl_chicago_retail = chicago_retail) %>%
  select(yw, gl_farm_price, gl_chicago_retail)

red_leaf_clust <- as_tibble(red_leaf) %>%
  mutate(rl_farm_price = farm_price, rl_chicago_retail = chicago_retail) %>%
  select(yw, rl_farm_price, rl_chicago_retail)

data2 <- join(romaine_clust, iceberg_clust, by = "yw", type = "full")
data3 <- join(data2, green_leaf_clust, by = "yw", type = "full")
data <- join(data3, red_leaf_clust, by = "yw", type = "full")


row.names(data) <- as.character(data$yw)
  
data <- data %>% 
  select(-yw)

data
```
Green leaf farm and retail prices actually have a lot of missing values in 2018. Rather than trying to interpolate, I'm goign to just drop those columns. Green leaf and red leaf prices look closely correlated, so the PCA would essentially take one out anyways.

There is at the end of 2017 a big spike in green leaf farm, not reflected in red leaf farm. Also in the middle of 2018 a big spike in red leaf retail not reflected in green leaf retail. Not sure what that's about. But again, I'm jsut dorpping green leaf data for the rest of this analysis

```{r}
data <- data %>% 
  select(-gl_farm_price, -gl_chicago_retail)

```



Interactive visualizations to play around with to see how the prices of different lettuce types relate to each other.

```{r}
library(plotly) 
```

```{r}
data_plot <- data%>%mutate(yw = row.names(data))

data_plot

p <- ggplot(data_plot, aes(x=log(ro_farm_price), y=log(rl_farm_price), 
                       color = format(as.Date(yearweek(yw)),format = "%Y"), 
                       text=yw)) +
  geom_point(alpha=0.75)

#ggplotly(p, tooltip="yw")
```

library for PCA and clustering visualizations to come.
```{r}
library(factoextra)
```


# PCA

There are still a few sporadic na values in the data. There's not many, so I will drop those rows.
```{r}
#drop rows containing any NA values
data_drop_na <- drop_na(data)
```

```{r}
#compute PCA
pca <- prcomp(data_drop_na, scale = TRUE)

#visualize eigenvalues
fviz_eig(pca)

#graph of observations
fviz_pca_ind(pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

#graph of variables
fviz_pca_var(pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

#biplot of observations and variables
fviz_pca_biplot(pca, repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )
```
Scree plot suggests we should keep the first 3 dimensions of the PCA - because there is an "elbow" in the scree plot there, and also because the first 3 components added up explain about 85% of the total variance.

I think maybe we should do 4 because it makes intuitive sense to me. The market is different for romaine/non-romaine lettuces during e. coli outbreaks, so we should need 4 dims to describe retail/farm prics for romainee/non-romaine lettuces.

```{r}
# First 4 principal components
comps <- data.frame(pca$x[,1:4])
```

We can make a cool 3D interactive plot
```{r}
#library(rgl)
# 3D plot (window pops out)

#plot3d(comps$PC1, comps$PC2, comps$PC3)
```

# K-Means clustering
```{r}
# Determine number of clusters
wss <- (nrow(comps)-1)*sum(apply(comps,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(comps, 
                                     nstart = 25, 
                                     iter.max=1000,
                                     centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")
```
I don't see an obvious elbow bend in the plot, so it's not clear to me how many clusters to do. I'll try a few possibilities. 

First, I'll do 2 clusters.

```{r}
set.seed(123) # set seed for reproducibility

#do the clustering
compsk2 <- kmeans(comps, 2, nstart = 25, iter.max=1000)

# 3D plot
#plot3d(comps$PC1, comps$PC2, comps$PC3, col=compsk2$clust)
```

```{r}
# Cluster sizes
sort(table(compsk2$clust))
clust <- names(sort(table(compsk2$clust)))
# First cluster
cluster1_of_2 = row.names(data[compsk2$clust==clust[1],])
# Second Cluster
cluster2_of_2 = row.names(data[compsk2$clust==clust[2],])

# cluster members
cluster1_of_2
```

So 2017 W12-16 and 2018 W45-46 were picked out as one cluster. Everything else was the other cluster.

```{r}
#visualize the clusters on a timeline

cluster1_of_2_df = data.frame(cluster1_of_2) %>%
  rename(yw = cluster1_of_2) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 1)

cluster2_of_2_df = data.frame(cluster2_of_2) %>%
  rename(yw = cluster2_of_2) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 2)

clusterk2_df = join(cluster1_of_2_df, cluster2_of_2_df, type="full") %>%
  mutate(cluster = as.character(cluster)) %>%
  as_tsibble(index = yw)

ggplot(clusterk2_df, aes(x=yw, y=0, color = cluster)) + geom_point()

unique(clusterk2_df$cluster)
```


Next, let's try 3 clusters.
```{r}
set.seed(123) # set seed for reproducibility

#do the clustering
compsk3 <- kmeans(comps, 3, nstart = 25, iter.max=1000)

# 3D plot
#plot3d(comps$PC1, comps$PC2, comps$PC3, col=compsk3$clust)
```

```{r}
# Cluster sizes
sort(table(compsk3$clust))
clust <- names(sort(table(compsk3$clust)))
# First cluster
cluster1_of_3 = row.names(data[compsk3$clust==clust[1],])
# Second Cluster
cluster2_of_3 = row.names(data[compsk3$clust==clust[2],])
# Third Cluster
cluster3_of_3 = row.names(data[compsk3$clust==clust[3],])
```

```{r fig.height=0.5, fig.width=3}
#visualize the clusters on a timeline

cluster1_of_3_df = data.frame(cluster1_of_3) %>%
  rename(yw = cluster1_of_3) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 1)

cluster2_of_3_df = data.frame(cluster2_of_3) %>%
  rename(yw = cluster2_of_3) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 2)

cluster3_of_3_df = data.frame(cluster3_of_3) %>%
  rename(yw = cluster3_of_3) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 3)

clusterk3_df_temp = join(cluster1_of_3_df, cluster2_of_3_df, type="full")
clusterk3_df = join(clusterk3_df_temp, cluster3_of_3_df, type = "full") %>%
  mutate(cluster = as.character(cluster)) %>%
  as_tsibble(index = yw)

ggplot(clusterk3_df) + geom_point(aes(x=yw, y=0, color = cluster))
```
Next, let's try 5 clusters.
```{r}
set.seed(123) # set seed for reproducibility

#do the clustering
compsk4 <- kmeans(comps, 5, nstart = 25, iter.max=1000)

# 3D plot
#plot3d(comps$PC1, comps$PC2, comps$PC3, col=compsk4$clust)
```

```{r}
# Cluster sizes
sort(table(compsk4$clust))
clust <- names(sort(table(compsk4$clust)))
# First cluster
cluster1_of_4 = row.names(data[compsk4$clust==clust[1],])
# Second Cluster
cluster2_of_4 = row.names(data[compsk4$clust==clust[2],])
# Third Cluster
cluster3_of_4 = row.names(data[compsk4$clust==clust[3],])
# Fourth Cluster
cluster4_of_4 = row.names(data[compsk4$clust==clust[4],])
# Fifth Cluster
cluster5_of_4 = row.names(data[compsk4$clust==clust[5],])
```

```{r fig.height=0.5, fig.width=3}
#visualize the clusters on a timeline

cluster1_of_4_df = data.frame(cluster1_of_4) %>%
  rename(yw = cluster1_of_4) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 1)

cluster2_of_4_df = data.frame(cluster2_of_4) %>%
  rename(yw = cluster2_of_4) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 2)

cluster3_of_4_df = data.frame(cluster3_of_4) %>%
  rename(yw = cluster3_of_4) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 3)

cluster4_of_4_df = data.frame(cluster4_of_4) %>%
  rename(yw = cluster4_of_4) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 4)

cluster5_of_4_df = data.frame(cluster5_of_4) %>%
  rename(yw = cluster5_of_4) %>%
  mutate(yw = yearweek(yw)) %>%
  mutate(cluster = 5)

clusterk4_df_temp = join(cluster1_of_4_df, cluster2_of_4_df, type="full")
clusterk4_df_temp_2 = join(clusterk4_df_temp, cluster3_of_4_df, type="full")
clusterk4_df_temp_3 = join(clusterk4_df_temp_2, cluster4_of_4_df, type="full")
clusterk4_df = join(clusterk4_df_temp_3, cluster5_of_4_df, type = "full") %>%
  mutate(cluster = as.character(cluster)) %>%
  as_tsibble(index = yw)

ggplot(clusterk4_df) + geom_point(aes(x=yw, y=0, color = cluster))
```

