---
output:
  pdf_document: default
  html_document: default
---
# Initial models for time series

#Read Data and Basic Plots
```{r}


#Load packages
library(readr)
library(fpp3) #this imports a few things including dplyr, tidyr,ggplot2, and more
library(tsibble)
library(forecast)
```

```{r}
#read in cvs as dataframe and convert to time series
#key = c(Mode,ORegionDAT, DRegionDAT) since these 4 categories each have one observation at each time
df <- readr::read_csv(file = "data_shipping_and_weather_joined_cleaned.csv") %>% mutate(yw = yearweek(yw)) %>% as_tsibble(key = c(Mode,ORegionDAT, DRegionDAT), index = yw)

head(df)
```
```{r}
autoplot(df, sanitized_cost)
```

```{r}
#seasonal plot for cost
gg_season(df, y = sanitized_cost, labels = "both") +
   labs(title = "Seasonal plot: Sanitized Cost")
```
```{r}
# seasonal plot for temperature
gg_season(df, y = tavg, labels = "both") +
   labs(title = "Seasonal plot: Avgerage daily temperature")
```

#Looking at Basic Models

Let's follow this article with our data set


```{r}
is.ts(df)
```

##Avg, Naive, SNaive (seasonal naive(!))

mean =  the forecasts of all future values are equal to the average (or “mean”) of the historical data

naive = the forecasts for every horizon correspond to the last observed value

Seasonal Naive = we set each forecast to be equal to the last observed value from the same season of the year


```{r eval = FALSE}
snaive(
  df,
  h = 2 * frequency(x),
  level = c(80, 95),
  fan = FALSE,
  lambda = NULL,
  biasadj = FALSE,
  ...,
  x = y
)
```

```{r}
#Look at just refridgerated trucks with dest IL_CHI
df_R_IL <- df %>%
  filter(df$Mode == "R", df$DRegionDAT == "IL_CHI")
```



```{r}
# Set training data from 2017 W01 to 2019 W27
train_R_IL <- df_R_IL %>%
  filter_index("2017 W01" ~ "2019 W27")
```

```{r}
# Fit the models
cost_fit <- train_R_IL %>%
  model(
    Mean = MEAN(sanitized_cost),
    `Naïve` = NAIVE(sanitized_cost),
    `Seasonal naïve` = SNAIVE(sanitized_cost)
  )
```

```{r}
# Generate forecasts for 54 weeks
cost_fc <- cost_fit %>% forecast(h = 100)
```

```{r}
# Plot forecasts against actual values
cost_fc %>%
  autoplot(df_R_IL, level = NULL) +
  labs(
    y = "Sanitized Cost",
    title = "Forecasts for weekly cost Refriderated to Chicago"
  ) +
  guides(colour = guide_legend(title = "Forecast"))

#mean =  the forecasts of all future values are equal to the average (or “mean”) of the historical data

#naive = the forecasts for every horizon correspond to the last observed value

#Seasonal Naive = we set each forecast to be equal to the last observed value from the same season of the year

  
```



```{r}
# Set training data from 2017 W01 to 2019 W27
train <- df %>%
  filter_index("2017 W01" ~ "2019 W27")
```

```{r}
# Fit the models
cost_fit <- train %>%
  model(
    Mean = MEAN(sanitized_cost),
    `Naïve` = NAIVE(sanitized_cost),
    `Seasonal naïve` = SNAIVE(sanitized_cost)
  )
```

```{r}
# Generate forecasts for 54 weeks
cost_fc <- cost_fit %>% forecast(h = 100)
```

```{r}
# Plot forecasts against actual values
cost_fc %>%
  autoplot(df,level = NULL) +
  labs(
    y = "Sanitized Cost",
    title = "Forecasts for weekly cost"
  )+
  guides(colour = guide_legend(title = "Forecast"))
```



