---
output:
  pdf_document: default
  html_document: default
---
#ARIMA/SARIMA

```{r}
library(plyr)
library(fpp3)
library(tsibble)
library(forecast)
library(zoo)
```

```{r}
#read in the interpolated data
data <- readr::read_csv(file = 'data/data_interpolated_with_lags.csv') %>%
  mutate(yw = yearweek(yw)) %>%
  select(-X1) %>%
  as_tsibble(key = c(Mode,ORegionDAT, DRegionDAT), index = yw) 
```

```{r}
#make into univariate approx_cost series
data <- data %>%
  select(Mode, ORegionDAT, DRegionDAT, yw, approx_cost, tmax_lag_12, tmax_lag_2, prcp_lag_12, prcp_lag_2, diesel_price, new_deaths, pandemic, volume) %>%
  filter(Mode == "R", DRegionDAT == "IL_CHI")
```

```{r}
#trim leading and trailing na's
data <- drop_na(data)
```

```{r}
#create training set - up through 2020 of the time series
train <- data %>% 
  filter_index(~ "2018 W52")
```

There are many possible arima models - based on choice of hyperparameters and whether to include seasonality or not. The ARIMA() function automatically chooses the best hyperparameters.
```{r}
fit <- train %>%
  model(ARIMA(approx_cost ~ tmax_lag_12 + tmax_lag_2 + prcp_lag_12 + prcp_lag_2 + diesel_price + new_deaths + pandemic + volume))
```
```{r}
#see what the automatically chosen arima models were.
report(fit)
```
It looks like for the first 3 time series ARIMA() automatically picked up on the period 52 seasonality. For Boston V data, it did not. Also it chose a different set of hyperparameters for each time series. 

```{r}
#forecast
#in order to produce the forecast of approx_cost, we need to feed in a forecast of tmax, prcp, and diesel_price. Instead of producing a forecast, I will just use the actual data here. 
future_data <- data %>% 
  filter_index("2019 W01"~"2019 W26") %>%
  select(Mode, ORegionDAT, DRegionDAT, yw, approx_cost, tmax_lag_12, tmax_lag_2, prcp_lag_12, prcp_lag_2, diesel_price, new_deaths, pandemic, volume)
fc <- fit %>% forecast(future_data)
```

```{r fig.height = 6, fig.width = 10}
#plot
fc %>%
  autoplot(train) +
  autolayer(data, colour = "black") +
  labs(title="SARIMA")
```

#Looking at fitted values and residuals

```{r}
#get fitted values and residuals
aug = augment(fit)
```

```{r fig.height = 3, fig.width = 8}
#autoplot them
autoplot(aug, .innov) +
  labs(title = "Residuals from the ARIMA/SARIMA model")
```

```{r}
#histograms
aug %>%
  ggplot(aes(x = .innov)) +
  geom_histogram() +
  facet_grid(rows = vars(Mode), cols = vars(DRegionDAT)) +
  labs(title = "Histograms of residuals")
```

```{r}
#acf
aug %>%
  ACF(.innov, lag_max = Inf) %>%
  autoplot() +
  facet_grid(rows = vars(Mode), cols = vars(DRegionDAT)) +
  labs(title = "ACF of Residuals")
```

## Looking at forecast errors

```{r}
accuracy(fc, data)
```